Active learning for efficient systematic reviews

### 2020-07-20 manuscript/manuscript/Ferdinands, G - MSBBSS.Rmd
corrctions July 20 2020

---
title: Active learning for efficient systematic reviews
subtitle: Evaluating models across research areas 
author: "Gerbrich Ferdinands"
# keywords:
header-includes:
    - \usepackage{setspace}\doublespacing
    - \usepackage[left]{lineno}
date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
  bookdown::pdf_document2:
    extra_dependencies: subfig, algorithm2e, float
    number_sections: no
    toc: no
  word_document: default
bibliography: asreview.bib
link-citations: yes
csl: elsevier-vancouver.csl
---

@@

<!--   bookdown::pdf_document2: -->
<!--     extra_dependencies: subfig, algorithm2e -->
<!--     number_sections: no -->
<!--     toc: no -->
<!--   word_document: default -->
<!-- bibliography: asreview.bib -->
<!-- link-citations: yes -->
<!-- csl: systematic-reviews.csl -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.ncol = 2, out.width = "50%", fig.align = "center", fig.pos = "H") 
options(tinytex.verbose = TRUE)

@@

```{r all tables, results = 'hide'}
table_nums <- captioner(prefix = "Table")
# table  on datasets statistics
table_nums(name = "datasets", caption = "Statistics on the test datasets obtained from six original systematic reviews.")
table_nums(name = "atd", caption = "ATD values ($\\bar x (\\hat s)$) for all model-dataset combinations, and median (MAD) for all datasets.")
table_nums(name = "wss", caption = "WSS@95 values ($\\bar x (\\hat s)$) for all model-dataset combinations, and median (MAD) for all datasets.")
table_nums(name = "rrf", caption = "RRF@10 values ($\\bar x, (\\hat s)$) for all model-dataset combinations, and median (MAD) for all datasets.")

@@

\linenumbers
# Background
Systematic reviews are top of the bill in research. A systematic review brings together all studies relevant to answer a specific research question [@PRISMA-PGroup2015]. Systematic reviews inform practice and policy [@Gough2002] and are key in developing clinical guidelines [@Chalmers2007]. However, systematic reviews are costly because they involve the manual screening of thousands of titles and abstracts to identify publications relevant to answering the research question.

<!-- An experienced reviewer takes on average 30 seconds to screen one title and abstract, whereas an inexperienced reviewer takes even longer [@Wallace2010]. A typical review of 5000 publications would require over 40 hours of uninterrupted screening, while typically only a fraction of the publications is relevant (mean = 2.94%, IQR=2.5 @Borah2017]). -->
Conducting a systematic review typically requires over a year of work by a team of researchers [@Borah2017]. Nevertheless, systematic reviewers are often bound to a limited budget and timeframe. Currently, the demand for systematic reviews exceeds the available time and resources by far [@Lau2019]. Especially when the need for guidelines is urgent it is almost impossible to provide a review that is both timely and comprehensive.

@@

A well-established approach in increasing efficiency in title and abstract screening is screening prioritization [@Cohen2009; @Shemilt2014]. In screening prioritization, the learning model presents the reviewer with the publications which are most likely to be relevant first, thereby expediting the process of finding all of the relevant publications. Such an approach allows for substantial time-savings in the screening process as the reviewer can decide to stop screening after a sufficient number of relevant publications have been retrieved [@Yu2019]. Moreover, reviewing relevant publications early facilitates a faster transition of those publications to the next steps in the review process [@Cohen2009]. 
<!-- Additionally, several studies report increasing efficiency beyond saving time [@OMara-Eves2015].  -->

<!-- assist the reviewer in the screening process -->
<!-- Several studies have proposed active learning models for screening prioritization -->
Recent studies have demonstrated the effectiveness of screening prioritization by means of active learning models [@Yu2018; @Yu2019; @Miwa2014; @Cormack2014; @Cormack2015; @Wallace2010; @Gates2018a]. With active learning, the machine learning model can iteratively improve its predictions on unlabelled data by allowing the model to select the records from which it wants to learn [@Settles2012]. The model queries these records to a human annotator who provides them with a label, from which the model then updates its predictions. The general assumption is that by letting the model select which records are labelled, the model can achieve higher accuracy while requiring the human annotator to label as few records as possible [@Settles2009]. Active learning has proven to be an efficient strategy in large unlabelled datasets where labels are expensive to obtain [@Settles2009]. This makes the screening phase in systematic reviewing an ideal candidate for such models because typically, labelling a large number of publications is very costly. When active learning is applied in the screening phase, the reviewer screens publications that are selected by an active learning model. Subsequently, the active learning model learns from the reviewers' decision ('relevant', 'irrelevant') and uses this knowledge to update its predictions and to select the next publication to be screened by the reviewer.

@@

The application of active learning models in reducing workload of systematic reviews has been extensively studied [@Yu2018; @Yu2019; @Miwa2014; @Wallace2010; @Gates2018a]. Whilst previous studies have evaluated active learning models in many forms and shapes [@Yu2018; @Yu2019; @Miwa2014; @Wallace2010], all studies used the same classification technique to predict relevance of publications, namely Support Vector Machine. Findings from outside the field of active learning show that different classification techniques can serve different needs in the retrieval of relevant publications, for example recall versus precision [@Kilicoglu2009; @Aphinyanaphongs2004]. Therefore, it is essential to evaluate different classification techniques in the context of active learning models.
<!-- no study in this area has made a comparison between models adopting different classification techniques. -->
<!-- Since a plethora of techniques exist [@Aggarwal2012], it would be of interest to incorporate such models to this area as well.  -->
Another component known to influence performance of the models is the way how the textual content of titles and abstracts are represented in a model, called the feature extraction strategy [@Le2014; @Zhang2011]. Previous studies all adopt an effective but rather simplistic 'bag of words' strategy [@Yu2018; @Yu2019; @Miwa2014; @Wallace2010]. It is of interest to evaluate models using this approach by comparing them to models adopting a more sophisticated strategy, called 'Doc2vec' [@Le2014]. Lastly, previous studies have mainly focussed on reviews from a single scientific field, like medicine [@Wallace2010; @Gates2018a] and software engineering [@Yu2018, @Yu2019]. Model replications on reviews from varying research contexts are essential to draw conclusions about the general effectiveness of active learning models [@OMara-Eves2015; @Marshall2020]. As far as known to the authors, Miwa et al [@Miwa2014] were the only researchers to make a direct comparison between two systematic reviews from different research areas, namely the social and the medical sciences. They found that active learning was more difficult on data from the social sciences due to the different nature of the vocabularies used. Therefore, it is of interest to evaluate model performance across different research contexts. Taken together, evaluations of active learning models in the context of systematic reviewing are required (1) across different classification techniques, (2) feature extraction strategies, and (3) review contexts. The current study aims to address these issues by answering the following research questions:

|    __RQ1__ What is the performance of several active learning models across different classification techniques?
|    __RQ2__ What is the performance of several active learning models across different feature extraction strategies? 
|    __RQ3__ Does the performance of active learning models differ across systematic reviews from different research areas?

The purpose of the current paper is to increase the evidence base on active learning models for reducing workload in title and abstract screening in systematic reviews. We adopt four different classification techniques (Naive Bayes, Linear Regression, Support Vector Machine, and Random Forest) and two different feature extraction strategies (TF-IDF and Doc2vec) for the purpose of maximizing the number of identified relevant publications, while minimizing the number of publications needed to screen. Model performance was assessed by conducting a simulation on six systematic review datasets. datasets were collected from the fields of medicine [@Cohen2006; @Appenzeller-Herzog2019], virology [@Kwok2020], software engineering [@Yu2018], behavioural public administration [@Nagtegaal2019] and psychology [@vandeSchoot2017], to assess generalizability of the models across research contexts. The models, datasets and simulations are implemented in a pipeline of active learning for screening prioritization, called $\texttt{ASReview}$ [@ASReview2020]. $\texttt{ASReview}$ is an open source and generic tool such that users can adapt and add modules as they like, encouraging fellow researchers to replicate findings from previous studies. All scripts and data used are openly published to facilitate usability and acceptability of ML-assisted title and abstract screening in the field of systematic review.

<!-- . -->

The remaining part of this paper is organized as follows. The Technical details section elaborates on the characteristics of active learning models for identifying relevant publications in the context of systematic reviews. The Simulation study section describes the study that was designed to answer the research questions. The findings of the simulation study are reported in the Results section. The implications of the findings in context of previous research are discussed in the Discussion section, followed by this study's main conclusions in the Conclusion section.

# Technical details 
What follows is a more detailed account of the active learning models. The structure and functions of the key components of the models are introduced to clarify the choices made in the design of the current study.

## Task description
The screening process of a systematic review starts with all publications obtained in the search. The task is to identify which of these publications are relevant, by screening them at the title and abstract level. In active learning for screening prioritization, the screening process proceeds as follows:

@@

- Start with the set of all unlabelled records (titles and abstracts), $\mathcal{U}$.
- The reviewer provides a label for a few, for example 5-10, records $x \in \mathcal{U}$, creating an set of labelled records $x \in \mathcal{L}$ such that $x \notin \mathcal{U}$. The label can be either Relevant $\langle x, \texttt{R}\rangle$ or Irrelevant $\langle x, \texttt{I}\rangle$.
- The active learning cycle starts:

  1. A classifier, $C$, is trained on the labelled records $\mathcal{L}$, $C = \texttt{train}(\mathcal{L})$

@@

  6. The newly labelled record is added to the training data, such that $x \in \mathcal{L}$ and $x \notin \mathcal{U}$

  7. Back to step 1.

In this active learning cycle, the model can incrementally improve its predictions on the remaining unlabelled title and abstracts. The relevant titles and abstracts are identified as early in the process as possible. The reviewer and the model keep interacting until the reviewer decides to stop or until all records been labelled.

## Class imbalance problem
There are two classes in the dataset: relevant and irrelevant publications. Typically, the inclusion rate is low as only a fraction of the publications belong to the relevant class (2.94%, [@Borah2017]). The class imbalance causes the classifier to miss relevant publications, because there are far fewer examples of relevant than irrelevant publications to train on [@OMara-Eves2015]. Moreover, classifiers can achieve high accuracy but still fail to identify any of the relevant publications [@Wallace2010]. This is evident in the case of a systematic review dataset where only three percent of publications are relevant. A model would achieve 97% accuracy when classifying all publications as irrelevant, even though none of the relevant papers would have been correctly identified.

Previous studies have addressed the class imbalance problem by rebalancing the training data in various ways [@OMara-Eves2015]. To decrease the class imbalance in the training data, the models in the current study rebalance the training set by Dynamic Resampling (DR). DR undersamples the number of irrelevant publications in the training data, whereas the number of relevant publications are oversampled such that the size of the training data remains the same. The ratio between relevant and irrelevant publications in the rebalanced training data is not fixed, but dynamically updated depending on the number of publications in the available training data, the number of publications in the total data, and the ratio between relevant and irrelevant publications in the available training data.

## Classification
To make predictions on the unlabelled publications, a classifier is trained on features from the training data. 
<!-- The classifier predicts class of a given unlabelled publication.  -->
A technique widely used in classification tasks is the Support Vector Machine (SVM). SVMs separate the data into classes by finding a multidimensional hyperplane [@Tong2001; @Kremer2014]. SVMs have been proven to be effective in active learning models for screening prioritization [@Yu2018; @Miwa2014]. Moreover, SVMs are the currently the only classifier implemented in ready-to-use software tools implementing active learning for screening prioritization (Abstrackr [@Wallace2012b], Colandr [@Cheng2018], FASTREAD [@Yu2018], Rayyan [@Ouzzani2016], and RobotAnalyst [@Przybyla2018]).

Whilst the performance of several classification techniques has been investigated in the ML-aided title-and-abstract screening field in general [@Kilicoglu2009; @Aphinyanaphongs2004], the relatively new subfield of active learning for screening prioritization has not yet studied the performance of classifiers other than SVMs [@Yu2018; @Yu2019; @Cormack2014; @Cormack2015; @Miwa2014; @Wallace2010]. 
<!-- Whilst the SVM is a well-known technique, also classifiers have been employed in class prediction tasks in systematic reviews.  -->
<!-- In fact, research in this area to date has not yet determined the performance of classifiers other than SVMs. -->
The current study aims to address this gap by exploring performance of three classifiers besides SVM:

 - L2-regularized Logistic Regression (LR) models the probabilities describing the possible outcomes by a logistic function. A penalty is imposed on the size of the coefficients, shrinking coefficients of features with a minor contribution to the solution towards zero.

 - Naive Bayes (NB) is a supervised learning algorithm often used in text classification. Based on Bayes' Theorem, with the 'naive' assumption that all features are independent given the class value [@Zhang2004].
- Random Forests (RF) is a supervised learning algorithm where a large number of decision trees are fit on bootstrapped samples of the original data. All trees cast a vote on the class, which are aggregated into a class prediction for each instance [@Breiman2001].

These three classification techniques were selected because they are widely adopted methods in text classification [@Aggarwal2012]. Moreover, these techniques can be run on a personal computer as they require a relatively low amount of processing power.

## Feature extraction
To predict publication class, the classifier uses information from the publications in the dataset. Examples of such information are titles and abstracts. However, a model cannot make predictions from the titles and abstracts as they are; their textual content needs to be represented numerically. The textual information needs to be mapped to feature vectors. This process of numerically representing textual content is referred to as 'feature extraction'.

A classical example of feature extraction is a 'bag of words' (bow) representation. For each text in the dataset, the term frequency - the number of occurrences of each word - is stored. This leads to $n$ features, where $n$ is the number of distinct words in the texts [@scikit-learn]. A serious weakness of this method is that it highly values often occurring but otherwise meaningless words such as "the". A more sophisticated bow approach is Term-Frequency Inverse Document Frequency (TF-IDF), which circumvents this problem by adjusting the term frequency in a text with the inverse document frequency, the frequency of a given word in the entire dataset [@Ramos2003]. A downside of TF-IDF and other bow methods is that they do not take into account the ordering of words, thereby ignoring semantics. An example of an approach that aims to overcome this weakness is Doc2vec (D2V). Doc2vec extracts features of the texts by a neural network, capable of grasping semantics by learning to predict the words in the texts [@Le2014].

## Query strategy
The active learning model can adopt different strategies in selecting the next publication to be screened by the reviewer. A strategy mentioned before is selecting the publication with the highest probability of being relevant. In the active learning literature this is referred to as certainty-based active learning [@Settles2012]. Another well-known strategy is uncertainty-based active learning, where the instances that are presented next are those instances on which the model's classifications are the least certain, i.e. close to 0.5 probability [@Settles2012]. Traditionally, this strategy trains the most accurate model because the model can learn the most from instances it is uncertain about. However, a study comparing performance of both strategies in detecting relevant publications found that the accuracy gain of uncertainty-based screening was not significant [@Miwa2014].

Certainty-based active learning is the preferred strategy for the task at hand. Firstly, this strategy is far better suited to the goal of prioritizing relevant publications compared to uncertainty-based active learning, in which the publications are prioritized that the model is most uncertain about. Secondly, certainty-based active learning is far better equipped at dealing with imbalanced data in active learning, as it aims to present only records that belong to the relevant class [@Fu2011].

# Simulation study
The section below describes the simulation study that was carried out to answer the research questions.

## Set-up
To address __RQ1__, four models combining every classifier with TF-IDF feature extraction were investigated:

@@

To address __RQ2__, the classifiers were combined with Doc2vec feature extraction, leading to the following three models:

@@

The combination NB + D2V could not be tested because the Multinomial Naive Bayes classifier^[https://scikit-learn.org/stable/modules/generated/sklearn.naive_bayes.MultinomialNB.html#sklearn.naive_bayes.MultinomialNB] can only handle a feature matrix with positive values, whereas the Doc2vec feature extraction approach^[https://radimrehurek.com/gensim/models/doc2vec.html] produces a feature matrix that can also contain negative values. Performance of the seven models was evaluated by simulating every model on six systematic review datasets, addressing __RQ3__. Hence, 42 simulations were carried out, representing all model-dataset combinations. 
Every simulation started with an initial training set of one relevant and one irrelevant publication to represent a 'worst case scenario' where the reviewer has minimal prior knowledge on the publications in the data. To account for sampling variance, every simulation was repeated for 15 trials, a reasonable number within the available computing power. To account for bias due to the content of the initial publications, the initial training set was randomly sampled from the dataset for every of the 15 trials. Although varying over trials, the 15 initial training sets were kept constant over datasets to allow for a direct comparison of models within datasets. A seed value was set to ensure reproducibility.For every simulation, hyperparameters were optimized through a random search to arrive at maximum model performance. The classifier was retrained every time after a publication had been labelled. The simulation ended after all publications in the dataset had been labelled. Simulations were run using \texttt{ASReview}'s simulation mode [@ASReview2020]. There was no need for a human reviewer as the model could query the labels in the data instead.

This study has been approved by the Ethics Committee of the Faculty of Social and Behavioural Sciences of Utrecht University, filed as an amendment under study 20-104. Simulations were run in ASReview, version 0.9.3 [@ASReview2020]. Analyses were carried out using \texttt{R}, version 3.6.1 [@RCoreTeam2019]. Scripts and data are stored in the GitHub repository for this thesis^[https://github.com/GerbrichFerdinands/asreview-thesis].The output resulting from the simulation was stored on the Open Science Framework page of this thesis ^[https://osf.io/7mr2g/], as their size exceeded the storage limit of GitHub by far. Due to their large number, the simulations were carried out on Cartesius, the Dutch national supercomputer. Access was granted by SURF via a grant (ID EINF-156).

>>>>

---
title: Active learning for efficient systematic reviews
subtitle: Evaluating models across research areas 
author: "Gerbrich Ferdinands"
# keywords:
header-includes:
    - \usepackage{setspace}\doublespacing
    - \usepackage[left]{lineno}
    - \usepackage{amssymb}
date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
  bookdown::pdf_document2:
    extra_dependencies: subfig, algorithm2e, float
    number_sections: no
    toc: no
  word_document: default
bibliography: asreview.bib
link-citations: yes
csl: elsevier-vancouver.csl
---

@@

<!--   bookdown::pdf_document2: -->
<!--     extra_dependencies: subfig, algorithm2e, float -->
<!--     number_sections: no -->
<!--     toc: no -->
<!--   word_document: default -->
<!-- bibliography: asreview.bib -->
<!-- link-citations: yes -->
<!-- csl: systematic-reviews.csl -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.ncol = 2, out.width = "50%", fig.align = "center", dev = "png", fig.pos = "H") 
options(tinytex.verbose = TRUE)

@@

```{r all tables, results = 'hide'}
table_nums <- captioner(prefix = "Table")
# table  on datasets statistics
table_nums(name = "datasets", caption = "Statistics on the datasets obtained from six original systematic reviews.")
table_nums(name = "atd", caption = "ATD values ($\\bar x (\\hat s)$) for all model-dataset combinations, and median (MAD) for all datasets.")
table_nums(name = "wss", caption = "WSS@95 values ($\\bar x (\\hat s)$) for all model-dataset combinations, and median (MAD) for all datasets.")
table_nums(name = "rrf", caption = "RRF@10 values ($\\bar x, (\\hat s)$) for all model-dataset combinations, and median (MAD) for all datasets.")

@@

\linenumbers
# Abstract
#### Background 
the context and purpose of the study

#### Methods
how the study was performed and statistical tests used

#### Results
the main findings

#### Conclusions
brief summary and potential implications

keywords:

# Background
Systematic reviews are top of the bill in research. A systematic review brings together all studies relevant to answer a specific research question [@PRISMA-PGroup2015]. Systematic reviews inform practice and policy [@Gough2002] and are key in developing clinical guidelines [@Chalmers2007]. However, systematic reviews are costly because they involve the manual screening of thousands of titles and abstracts to identify publications relevant to answering the research question.

<!-- An experienced reviewer takes on average 30 seconds to screen one title and abstract, whereas an inexperienced reviewer takes even longer [@Wallace2010]. A typical review of 5000 publications would require over 40 hours of uninterrupted screening, while typically only a fraction of the publications is relevant (mean = 2.94%, IQR=2.5 @Borah2017]). -->
Conducting a systematic review typically requires over a year of work by a team of researchers [@Borah2017]. Nevertheless, systematic reviewers are often bound to a limited budget and timeframe. Currently, the demand for systematic reviews exceeds the available time and resources by far [@Lau2019]. Especially when the need for guidelines is urgent it is extremely challenging to provide a review that is both timely and comprehensive.

@@

A well-established approach in increasing efficiency in title and abstract screening is screening prioritization [@Cohen2009; @Shemilt2014]. In screening prioritization, the learning model presents the reviewer with the publications that are most likely to be relevant first, thereby expediting the process of finding all of the relevant publications. Such an approach allows for substantial time-savings in the screening process as the reviewer can decide to stop screening after a sufficient number of relevant publications have been retrieved [@Yu2019]. Moreover, the early retrieval of relevant publications facilitates a faster transition of those publications to the next steps in the review process [@Cohen2009]. 
<!-- Additionally, several studies report increasing efficiency beyond saving time [@OMara-Eves2015].  -->

<!-- assist the reviewer in the screening process -->
<!-- Several studies have proposed active learning models for screening prioritization -->
Recent studies have demonstrated the effectiveness of screening prioritization by means of active learning models [@Yu2018; @Yu2019; @Miwa2014; @Cormack2014; @Cormack2015; @Wallace2010; @Gates2018a]. With active learning, the ML model can iteratively improve its predictions on unlabelled data by allowing the model to select the records from which it wants to learn [@Settles2012]. The model queries these records to a human annotator who provides them with a label, from which the model then updates its predictions. The general assumption is that by letting the model select which records are labelled, the model can achieve higher accuracy more quickly while requiring the human annotator to label as few records as possible [@Settles2009]. Active learning has proven to be an efficient strategy in large unlabelled datasets where labels are expensive to obtain [@Settles2009]. This makes the screening phase in systematic reviewing an ideal candidate solution for such models because typically, labelling a large number of publications is very costly. When active learning is applied in the screening phase, the reviewer screens publications that are suggested by an active learning model. Subsequently, the active learning model learns from the reviewers' decision ('relevant', 'irrelevant') and uses this knowledge to update its predictions and to select the next publication to be screened by the reviewer.

@@

The application of active learning models in reducing workload of systematic reviews has been extensively studied [@Yu2018; @Yu2019; @Miwa2014; @Wallace2010; @Gates2018a]. Whilst previous studies have evaluated active learning models in many forms and shapes [@Yu2018; @Yu2019; @Miwa2014; @Cormack2014; @Cormack2015; @Wallace2010; @Singh2018; @Carvallo; @Yimin2007], ready-to-use software tools implementing such active learning models (Abstrackr [@Wallace2012b], Colandr [@Cheng2018], FASTREAD [@Yu2018], Rayyan [@Ouzzani2016], and RobotAnalyst [@Przybyla2018]) currently use the same classification technique to predict relevance of publications, namely Support Vector Machine (SVM). Findings show that different classification techniques can serve different needs in the retrieval of relevant publications, for example the desired balance between recall and precision [@Kilicoglu2009; @Aphinyanaphongs2004]. Therefore, is essential to evaluate different classification techniques in the context of active learning models. The current study investigates active learning models adopting four classification techniques: Naive Bayes, Logistic Regression, Support Vector Machine, and Random Forest. These are widely adopted techniques in text classification [@Aggarwal2012] and are suitable for creating interpretable ML models [@Molnar2020]. Moreover, due to their relatively short computation time they are fit for software tools to be used in scientific practice, as they allow the reviewer to be timely presented with the next publication.

<!-- low computation time makes these techniques fit to practial use in a software tool when run on a personal computer, as retraining a model can be done in a timely manner. as default of asreview is on local computer to be applicable -->

<!-- Whilst the performance of several classification techniques has been investigated in the ML-aided title-and-abstract screening field in general [@Kilicoglu2009; @Aphinyanaphongs2004], the relatively new subfield of active learning for screening prioritization has not yet studied the performance of classifiers other than SVMs [@Yu2018; @Yu2019; @Cormack2014; @Cormack2015; @Miwa2014; @Wallace2010]. -->

<!-- no study in this area has made a comparison between models adopting different classification techniques. -->
<!-- Since a plethora of techniques exist [@Aggarwal2012], it would be of interest to incorporate such models to this area as well.  -->
Another component that influences model performance is the way how the textual content of titles and abstracts are represented in a model, called the feature extraction strategy [@Zhang2011; @Carvallo; @Singh2018]. One of the more sophisticated feature extraction strategies is called Doc2vec (D2V), also known as Paragraph vector [@Le2014]. D2V learns continuous distributed vector representations for pieces of text. The idea of distributed representation is that every word can be known by its neighborhood. D2V can overcome problems with more simplistic feature extraction strategies such as TF-IDF, which ignores the the semantics and the ordering of the words [@Le2014]. Therefore, it is of interest to compare models adopting D2V to models adopting TF-IDF.

<!-- [@Yu2018; @Yu2019; @Miwa2014; @Wallace2010], -->

Lastly, previous studies have mainly focussed on reviews from a single scientific field, like medicine [@Wallace2010; @Gates2018a] and computer science [@Yu2018, @Yu2019]. Model replications on reviews from varying research contexts are essential to draw conclusions about the general effectiveness of active learning models [@OMara-Eves2015; @Marshall2020]. As far as known to the authors, Miwa et al [@Miwa2014] were the only researchers to make a direct comparison between two systematic reviews from different research areas, namely the social and the medical sciences. They found that active learning was more difficult on data from the social sciences due to the different nature of the vocabularies used. Thus, it is of interest to evaluate model performance across different research contexts, namely the social science, humanities, medical science and computer science.

Taken together, for a more comprehensive understanding of active learning models in the context of systematic reviewing, a methodical evaluation of such models is required. The current study aims to address this issue by answering the following research questions:

<!-- (1) across different classification techniques, (2) feature extraction strategies, and (3) review contexts.  -->

<!-- Therefore, the current study aims to address these issues by answering the following research questions: -->

|    __RQ1__ What is the performance of active learning models across four different classification techniques?
|    __RQ2__ What is the performance of active learning models across two different feature extraction strategies? 
|    __RQ3__ Does the performance of active learning models differ across six systematic reviews from four different research areas?

The purpose of the current paper is to increase the evidence base on active learning models for reducing workload in title and abstract screening in systematic reviews. We adopt four different classification techniques (NB, LR, SVM, and RF) and two different feature extraction strategies (TF-IDF and D2V) for the purpose of maximizing the number of identified relevant publications, while minimizing the number of publications needed to screen. Models were assessed by conducting a simulation on six systematic review datasets. Datasets were collected from the fields of medical science [@Cohen2006; @Appenzeller-Herzog2019; @Kwok2020], computer science [@Yu2018], humanities [@Nagtegaal2019] and social science [@vandeSchoot2017], to assess generalizability of the models across research contexts. The models, datasets and simulations are implemented in a pipeline of active learning for screening prioritization, called $\texttt{ASReview}$ [@vandeSchoot2020]. $\texttt{ASReview}$ is an open source and generic tool in which users can adapt and add modules as they like, encouraging fellow researchers to replicate findings from previous studies. All scripts and data used are openly published to facilitate usability and acceptability of ML-assisted title and abstract screening in the field of systematic review.

The remaining part of this paper is organized as follows. The methods section describes the study that was designed to answer the research questions and elaborates on the technical details of active learning models for identifying relevant publications in the context of systematic reviews. The findings of the simulation study are reported in the results section. The implications of the findings in context of previous research are discussed in the discussion section, followed by this study's main conclusions in the conclusion section.

# Methods

## Technical details 
What follows is a more detailed account of the active learning models. The structure and functions of the key components of the models are introduced to clarify the choices made in the design of the current study.

### Task description
The screening process of a systematic review starts with all publications obtained in the search. The task is to identify which of these publications are relevant, by screening their titles and abstracts. In active learning for screening prioritization, the screening process proceeds as follows:

@@

- Start with the set of all unlabelled records (titles and abstracts), $\mathcal{U}$.
- The reviewer provides a label for a few, for example 5-10 records $x \in \mathcal{U}$, creating a set of labelled records $x \in \mathcal{L}$ such that $x \notin \mathcal{U}$. The label can be either Relevant $\langle x, \texttt{R}\rangle$ or Irrelevant $\langle x, \texttt{I}\rangle$.
- The active learning cycle starts:

  1. A classifier $C$ is trained on the labelled records $\mathcal{L}$, $C = \texttt{train}(\mathcal{L})$

@@

  6. The newly labelled record is added to the training data, such that $x^* \in \mathcal{L}$ and $x^* \notin \mathcal{U}$

  7. Back to step 1

  8. Repeat this cycle until the reviewer decides to stop [@Yu2019] or until all records have been labelled, $\mathcal{U} = \varnothing$.

In this active learning cycle, the model incrementally improves its predictions on the remaining unlabelled title and abstracts. Relevant titles and abstracts are identified as early in the process as possible.

This case is an example of pool-based active learning, as the next record to be queried is selected by predicting relevancy for all records in a fixed pool [@Settles2012]. Another form of active learning is stream-based active learning, in which data is regarded as a stream instead of a fixed pool, in which the model selects one record at a time and then decides whether or not to query this record. This approach of active learning is preferred when it is expensive or impossible to exhaustively search the data for selecting the next query. A possible application of stream-based active learning is living systematic reviews, as the review is continually updated as new evidence becomes available. For an example see the study by Wynants et al., @Wynants2020.

### Class imbalance problem
There are two classes in the dataset: relevant and irrelevant publications. Typically, the inclusion rate is low as only a fraction of the publications belong to the relevant class (2.94%, [@Borah2017]). The class imbalance causes the classifier to miss relevant publications, because there are far fewer examples of relevant than irrelevant publications to train on [@OMara-Eves2015]. Moreover, classifiers can achieve high accuracy but still fail to identify any of the relevant publications [@Wallace2010]. This is evident in the case of a systematic review dataset where only three percent of publications are relevant.

<!-- A model would achieve 97% accuracy when classifying all publications as irrelevant, even though none of the relevant papers would have been correctly identified.  -->

Previous studies have addressed the class imbalance problem by rebalancing the training data in various ways [@OMara-Eves2015]. To decrease the class imbalance in the training data, the models in the current study rebalance the training set by a technique we propose to call Dynamic Resampling (DR). DR undersamples the number of irrelevant publications in the training data, whereas the number of relevant publications are oversampled such that the size of the training data remains the same. The ratio between relevant and irrelevant publications in the rebalanced training data is not fixed, but dynamically updated depending on the number of publications in the available training data, the total number of publications in the dataset, and the ratio between relevant and irrelevant publications in the available training data.

<!-- This works because..  -->
<!-- This technique is related to aggressive undersampling in such a way that.  -->

### Classification
To make predictions on the unlabelled publications, a classifier is trained on features from the training data. The performance of the following four classifiers is explored:
<!-- The classifier predicts class of a given unlabelled publication.  -->
<!-- Whilst the SVM is a well-known technique, also classifiers have been employed in class prediction tasks in systematic reviews.  -->
<!-- In fact, research in this area to date has not yet determined the performance of classifiers other than SVMs. -->

 - Support Vector Machine (SVM) - SVMs separate the data into classes by finding a multidimensional hyperplane [@Tong2001; @Kremer2014].

 - L2-regularized Logistic Regression (LR) models the probabilities describing the possible outcomes by a logistic function. The coefficients are regularizes/model is regularized. 
 A penalty is imposed on the size of the coefficients, shrinking coefficients of features with a minor contribution to the solution towards zero.

 - Naive Bayes (NB) is a supervised learning algorithm often used in text classification. Based on Bayes' theorem, with the 'naive' assumption that all features are independent given the class value [@Zhang2004].

- Random Forests (RF) is a supervised learning algorithm where a large number of decision trees are fit on bootstrapped samples of the original data. All trees cast a vote on the class, which are aggregated into a class prediction for each instance [@Breiman2001].

### Feature extraction
To predict relevance of a given publication, the classifier uses information from the publications in the dataset. Examples of such information are titles and abstracts. However, a model cannot make predictions from the titles and abstracts as they are; their textual content needs to be represented numerically. The textual information needs to be mapped to feature vectors. This process of numerically representing textual content is referred to as 'feature extraction'.

Term-Frequency Inverse Document Frequency (TF-IDF) stores the term frequency - the number of occurrences of each word - for each text in the dataset. The term frequency is adjusted with the inverse document frequency, the number of documents which contain the given word [@Ramos2003]. A disadvantage of TF-IDF and other bag-of-words methods is that they do not take the ordering of words into account, thereby ignoring semantics. An example of an approach that aims to overcome this weakness is D2V. D2V learns continuous distributed vector representations for pieces of text capable of grasping semantics by learning to predict the words in the texts [@Le2014].

### Query strategy
The active learning model can adopt different strategies in selecting the next publication to be screened by the reviewer. A strategy mentioned before is selecting the publication with the highest probability of being relevant. In the active learning literature this is referred to as certainty-based active learning [@Settles2012]. Another well-known strategy is uncertainty-based active learning, where the instances that are presented next are those instances on which the model's classifications are the least certain, i.e. close to 0.5 probability [@Settles2012]. Traditionally, this strategy trains the most accurate model because the model can learn the most from instances it is uncertain about. However, a study comparing the performance of both strategies in detecting relevant publications found that the accuracy gain of uncertainty-based screening was not significant [@Miwa2014].

Certainty-based active learning is the preferred strategy for the task at hand. Firstly, the aim of the screening process is to find relevant publications, not to train a good model. Secondly, certainty-based active learning is far better equipped at dealing with imbalanced data in active learning, as it aims to present only records that belong to the relevant class [@Fu2011].

## Simulation study
The section below describes the simulation study that was carried out to answer the research questions.

### Set-up
To address __RQ1__, four models combining every classifier with TF-IDF feature extraction were investigated:

@@

To address __RQ2__, the classifiers were combined with D2V feature extraction, leading to the following three models:

@@

The combination NB + D2V could not be tested because the multinomial naive Bayes classifier^[https://scikit-learn.org/stable/modules/generated/sklearn.naive_bayes.MultinomialNB.html#sklearn.naive_bayes.MultinomialNB] can only handle a feature matrix with positive values, whereas the D2V feature extraction approach^[https://radimrehurek.com/gensim/models/doc2vec.html] produces a feature matrix that can also contain negative values. The performance of the seven models was evaluated by simulating every model on six systematic review datasets, addressing __RQ3__. Hence, 42 simulations were carried out, representing all model-dataset combinations.

Instead of having a human reviewer label publications manually, the screening process was simulated by retrieving the labels in the data. Each simulation started with an initial training set of one relevant and one irrelevant publication to represent a 'worst case scenario' where the reviewer has minimal prior knowledge on the publications in the data. The model was retrained every time after a publication had been labelled. A simulation ended after all publications in the dataset had been labelled. To account for sampling variance, each simulation was repeated 15 times. To account for bias due to the content of the initial publications, the initial training set was randomly sampled from the dataset for each of the 15 trials. Although varying over trials, the 15 initial training sets were kept constant over datasets to allow for a direct comparison of models within datasets. A seed value was set to ensure reproducibility. The simulation study was carried out using the \texttt{ASReview} simulation extension [@ASReview2020]. For each simulation, hyperparameters were optimized through a ... tpe to arrive at maximum model performance.  todo
<!-- About a 1000 runs, retraining after labelling 20, after visual inspection.  --> 
Scripts for the hyperparaemter optimization and resulting hyperparameter values can be found at GitHub.

Simulations were carried out in ASReview, version 0.9.3 [@ASReview2020]. Analyses were carried out using \texttt{R}, version 3.6.1 [@RCoreTeam2019]. Scripts and data are stored in the GitHub repository for this paper^[https://github.com/asreview/paper-evaluating-models-across-research-areas].The output resulting from the simulation was stored on the Open Science Framework page of this paper ^[https://osf.io/7mr2g/], as their size exceeded the storage limit of GitHub by far. Due to their large number, the simulations were carried out on Cartesius, the Dutch national supercomputer. Access was granted by SURF via a grant (ID EINF-156).

### Datasets 
The models were simulated on a convenience sample of six systematic review datasets. The data selection process was driven by two factors. Firstly, datasets were collected from various research areas to assess generalizability of the models across research contexts (RQ3). Secondly, all original data files should be openly published with a CC-BY license.
